Hysteresis in human computation:\\ how one task affects another

### 2015-01-22 docs/publications/2014.09_SCIENCE/man.tex
revisions to manuscript -- lots of work on supplementary material

\begin{sciabstract}

Faced with the ever-growing scale and complexity of data, academic and 
industrial researchers are increasingly relying on \textit{human computation} 
to extract useful knowledge.  Microtask
platforms, like Amazon Mechanical Turk, act like a human compute cluster,
making it possible to access human cognition quickly and inexpensively. 
This has lead to their broad adoption by social and computer scientists.  
Human computation is being used experimentally in applications
where expert opinion would normally be sought, such as in reviewing 
medical images.
Here we show that the repeated performance of similar tasks, an inherent
feature of microtask and crowdsourcing platforms, acts as a strong
and largely unexplored source of bias.  The effects that earlier tasks exert
on later ones are more severe than effects caused by framing, 
such as by disclosing a task's purpose.  
Using image-labeling, the canonical microtask, we show that prior 
tasks influence the content that workers chose to focus on, as well as the 
richness and specialization of their vocabulary.    
While intertask effects can be a source of systematic bias, our
results suggest that with appropriate task-engineering, it might be possible
to leverage intertask effects to hone worker focus and specificity, helping
to elicit reproducible, expert-level judgments.
\end{sciabstract}

@@

These platforms allow requesters to make use of human intelligence
\textit{algorithmically}.  Tasks are posted and results fetched by transfering
files over the Internet, making it possible to embed microtask work in an
automated process.  To the surrounding process, workers seem like 
input-output devices, which provides much of the flexibility and 
cost-savings of actually automating the work: the workforce is available
on demand, over the Internet, without any contracts or interviews, at a rate
of about 6 USD\$/ hr.  Many kinds of work can be performed at a fraction of 
the cost that would otherwise be required, either to recruit temporary 
workers, or to research and design a sophisticated automated solution.

@@

Given the increasing prevalence of microtask-based methods, there
is a desire to understand the factors affecting the reliability of microtask
platforms.  Researchers have investigated task-design parameters such as the 
level of pay \cite{kazai2013analysis}, the training and pre-screening of 
workers \cite{le2010ensuring,paolacci2010running}, and user-interface 
design \cite{Finnerty2013}.  
Researchers have also investigated the effects of \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.

Here we focus on a ubiquitous feature of the microtask setup: the 
tendency for microtask workers to perform many tasks in quick succession. 
Even if a worker performs the minimum possible amount of work, 
she will likely have performed several microtasks, since they are typically 
distributed to workers in bundles.  For example, in one bundle, a worker 
might be assigned a set of ten images, and asked to provide descriptive words 
(or \textit{labels}) for each.  On MTurk, these bundles are called Human 
Intelligence Tasks (HITs), but for the sake of clarity, we will use 
``task'' to mean the smallest repeatable unit of work, which, in our study, 
consists of labeling a single image.

@@

Researchers investigating a somewhat similar class of psychological phenomena,
found that ratings that microtask workers give to items in a list
depend on the order in which the list is presented, an effect known as
\textit{positional bias} \cite{lerman2014leveraging}.  This serves as an 
example of a basic psychological phenomenon impacting crowdsourced data,
and raises a serious question about what other cognitive biases might
affect microtask work, and how prevelant induced bias might be.

We therefore turn to the question of bias induced by taks-repetition effects, 
which, if present, would impact virtually every microtask workflow.
The results we present here show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
We call the effects that earlier tasks exert on later ones 
\textit{intertask effects}.

As we have described, microtask platforms have been widely adopted as a 
tool, but computer scientists also hold such platforms as the object of 
research, viewing  microtask platforms as a new kind of computing 
architecture.  In analogy to CPUs (central 
processing units) and GPUs (graphics processing units), 
Davis \textit{et al} invoked the \textit{human processing unit}, or HPU \cite{5543192}.
Researchers sought to define a basic HPU instruction set, and
libraries for HPU programming are under active development
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
HPU outputs are inherently stochastic, but our revelation of strong intertask 
effects shows that HPUs are subject to \textit{hysteresis}, 
meaning that their 
outputs depend on the history of their inputs.  Reproducibility and 
path-independance are crucial for properties for the basic components in a 
computing architecture, so it will be necessary to understand the 
computational ramifications of these psychological phenomena to further this 
interesting line of investigation.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  
In addition to being one of the most common kinds of tasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running},
image-labeling is prototypical for any kind of task in which the worker is 
asked to provide a qualitative free-text response to a prompt.  
This includes such such examples as describing
video and audio recordings, summarizing texts, or providing personal 
experiential feedback to a product review.
%in the long run, because image labeling is still very difficult to automate
%\cite{5543192}, and is used in many applications, such as website moderation, 
%computer vision research, and social media analytics (\# insert other refs) 
%\cite{5543192}.

In our first experiment, which we will refer to as 
\textit{intertask-food-objects}, workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in these treatments
performed a set of five \textit{initial tasks}, which each involved 
providing five descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (see Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  The test tasks contained images depicting a 
mixture of food and objects (see Fig.~\ref{fig:task}C), and workers
were again asked to provide five labels.  Five initial 
tasks, together with the test tasks, comprised one HIT 
(\# workers were only allowed to participate once). During the HIT, workers 
were presented images one at a time, and, crucially, we made no 
distinction between the initial and test tasks.

@@

While this affirms the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced by prior task 
exposures.  To this end, we define the extent of bias, $\theta$, in terms of
the strength of influence that a perturbation has on worker's responses.
Since different workers generally provide different responses for the same 
task, we characterize worker's responses by a probability distribution.  
The extent of bias is then defined by the difference in the response 
distribution for different treatments of workers
(\# this is a general divergence metric between probability 
distributions known as the \textit{total variation distance}):
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
	\label{eq:theta}
\end{equation}
where $p_i(x)$ is the probability that a worker subjected to treatment $i$
provides response $x$, from the set of possible responses $X$.
It is important to emphasize that a significant result from a $\chi^2$ test 
only allows us to conclude that $\theta$ is non-zero.  Determining $\theta$ from finite samples of 
responses, using purely statistical approaches, is difficult (\#), which 
leads us to reformulate the
problem in terms of classification: given a worker's 
response, can a machine reliably infer what her prior task-exposure had 
been?  Intuitively, the greater the effect that prior tasks have on subsequent
responses, the more accurate the classifier can be.  Formally, as we show
in the supplementary text, the accuracy of such a classifier, $\eta$,
yields a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
By measuring the accuracy of a classifier, we can bound the 
bias induced by any perturbation, including both intertask effects and 
framing.  As defined, bias ranges from 0 (meaning the exposures 
do not affect the probabilities of responses), to 1 
(or 100\%) (meaning that under different exposures, workers provide completely
different sets of responses).

@@

Using a na\"ive Bayes classifier (\# in the 
supplementary material we discuss the choice of classifier, the
approach to measuring accuracy, and draw a comparison to the results obtained 
if a support vector machine is used), we found that intertask effects 
lead to a 30\% bias between workers from the \textit{food} and 
\textit{objects} treatments.  This represents a substantial potential to 
distort microtask data.

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we performed a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias we found was statistically 
insignificant (\td{insert p-value})
(see Fig.~\ref{fig:theta}).  This suggests 
that prior task exposure is a stronger priming modality than framing.

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
As before, the initial tasks for the \textit{food} treatment contained images 
depicting food (see Fig.~\ref{fig:task}D). 
In the \textit{culture} treatment, 
the initial tasks contained images depicting cultural scenes 
(of dance, sport, or music) (see Fig.~\ref{fig:task}E).  The test tasks, which were identical for both
treatments, included images depicting meals of identifiable cultural 
origin (\# the initial tasks in 
\textit{intertask-food-culture} are arguably also of identifiable cultural 
origin, but this would tend to reduce, rather than inflate, our results
(we discuss this further in the supplementary material).)
(see Fig.~\ref{fig:task}F).

Results for this experiment again showed a 
strong bias as a result of intertask effects (about 50\%) 
(see Fig.~\ref{fig:theta}A).  Again, we compared intertask effects to 
framing, by performing an experiment (\textit{frame-food-culture}) using 
the same test tasks as in \textit{intertask-food-culture}, in which we exposed
workers to a statement framing the purpose of the tasks as the recognition 
of either food or culture.  
For this framing experiment, we did not observe a statistically significant 
difference between the labels attributed by the \textit{food} and 
\textit{culture} treatments during test tasks (see Fig.~\ref{fig:theta}).

@@

To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, during
the test tasks within
a given experiment, the food-primed workers would use food-related words 
more often than their non-food-primed counterparts.  However, this was not 
generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks
(\# food related words were identified using the wordnet corpus, augmented
with additional words obtained by crawling a recipe website (see supplementary
information for details)) (see Fig.~\ref{fig:specificity}).  This finding
rules out a seemingly-simple idea that workers emphasize
content that has been present in earlier tasks: seeing content in prior tasks
affects the probability that a worker refers to it in subsequent tasks, but 
does not necessarily \textit{increase} the probability.

To deepen our understanding, we investigated workers' lexical richness in 
reference to food, that is, the number of \textit{unique} food-related words
used.  Even if workers provide an abundance of food-related words, if the 
same words are often repeated, there will be less diversity.  
This could happen, for example, if workers use generic references 
to food.
Both \textit{intertask} experiments showed that food-exposed workers had 
greater lexical richness, in reference to food, than their non-food-primed 
counterparts (as much as 20\% more) (see Fig.~\ref{fig:specificity}).  
The increase in lexical richness in the case of 
\textit{intertask-food-culture} is particularly noteworthy, because in that 
experiment, food-primed workers made fewer total references to food.  
We also observed enrichment of the food-related lexicon in the 
\textit{priming-food-objects}, and \textit{echo-food-objects} experiments, 
although to a lesser extent.

The observations regarding lexical richness lead us to suspect that 
initial tasks might influence workers to use more refined or specialized 
words, when referring to aspects of content that had been present in the 
initial tasks (i.e. food).  This would help explain why, in the case of 
\textit{intertask-food-culture}, we observed a significantly greater 
\textit{diversity} of food-related words despite their significantly lesser 
\textit{total amount}.
To test whether exposure to food in initial tasks increased the specialization
of food-references, we used the wordnet corpus to operationalize the 
notion of word specialization.  Wordnet provides a set of hypernym-hyponym 
relations between \# English nouns.  Hypernyms are generalizations 
(for example, ``bread'' is a hypernym for 
``pumpernickel''), while hyponyms are specializations.
For each experiment, we compared food-related words provided by workers from 
one treatment to those from the other, and tallied the excess number of cases
where one treatment's words were more specialized than the other, as a 
percentage of the total (see Fig.~\ref{fig:specificity}B) (\# full detail for 
this calculation is provided the supplementary text).

In all experiments except \textit{frame-food-culture}, food-primed workers 
used significantly more specialized words than their non-food-primed 
counterparts (about 15\% more; see Fig.~\ref{fig:specificity}C).
It is interesting that such substantial increases in both the lexical 
richness and specialization of food-related words 
held for \textit{intertask-food-culture}, where, as mentioned, we observed 
that food-exposed workers made \textit{fewer} references to food overall. 
These observations point 
to countervailing factors: one factor tending to activate the more 
specialized and less common food-related words 
(yielding greater lexical richness and specialization), and the other tending 
to suppress certain, presumably more common and generic words 
(yielding fewer food-related words in total).

@@

\section*{Supplementary Material}

\subsection*{S1: Experimental materials}
The tasks for both experiments were presented as a series of slides.  In
both cases, the first slide consisted of a brief set of instructions, followed
by the frame if shown (only some treatments involved a frame), see 
Fig. S\ref{fig:hit_preamble}.  Following the instructions and prime, a set
of initial tasks was shown (except for the treatments in \textit{frame2}
and \textit{echo1}) followed by (in all cases) a set of test tasks.  In
\textit{exp1}, two kinds of frame were used, one simple frame, and one where
the worker was required to echo the frame using a combo-box input.  The
instructions, both kinds of frame, and an example task slide, are shown 
in Fig. S1.
\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\label{fig:hit_preamble}
	\caption{Examples of A) instructions; B) a frame, as shown in 
		\textit{frame1:obj}, and similar to those shown in 
		\textit{frame1:food} and \textit{frame2:food}, and 
		\textit{frame2:cult};  
		C) an echoed frame, as used in \textit{echo1:obj}, and similar to that
		used in \textit{echo1:food}; and D) an 
		example of an image-labeling task.
	}
\end{figure}

All of the initial tasks had the format shown in Fig. S1D.  The images used
in the initial tasks for \textit{task1:food} and \textit{task1:obj} are
shown in Fig. S2 and S3, and the images used in the test tasks for 
\textit{exp1} are shown in Fig. S4.  The images used in the initial tasks
for \textit{task2:food} and \textit{task2:cult} are shown in Fig. S5 and
Fig. S6 respectively, and the images used in the test tasks for \textit{exp2} 
are shown in Fig. S7.

\begin{figure}
	\includegraphics{figs/task1-food.pdf}
	\label{fig:task1:food}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task1:food}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-obj.pdf}
	\label{fig:task1:obj}
	\caption{
		Figure S3: Images used in the initial tasks for 
		\textit{task1:obj}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-test.pdf}
	\label{fig:task1:test}
	\caption{
		Figure S2: Images used in the test tasks for 
		\textit{exp1}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-food.pdf}
	\label{fig:task2:food}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task2:food}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-cult.pdf}
	\label{fig:task2:cult}
	\caption{
		Figure S2: Images used in the initial tasks for 
		\textit{task2:cult}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\begin{figure}
	\includegraphics{figs/task2-test.pdf}
	\label{fig:task2:test}
	\caption{
		Figure S2: Images used in the test tasks for 
		\textit{exp2}.  The numbers show the order in which the 
		images were presented to workers.
	}
\end{figure}

\subsection*{S2: Access to raw data and code}
The raw data and source code needed to reproduce the analyses 
shown here and in the main text can be downloaded from 
http://networkdynamics.org/publications/public/priming2014.09.

>>>>

\begin{sciabstract}
Microtask platforms have become commonplace tools for obtaining human 
annotations for large datasets.  Such platforms connect requesters 
(researchers or companies) with large populations (crowds) of workers, who 
will individually complete small information-processing jobs, each taking 
typically between one to five minutes.  
A major challenge in using such platforms, 
and a topic of ongoing research, concerns designing jobs that elicit high 
quality annotations. Here we identify a feature of nearly all crowdsourcing 
jobs that can act as a strong and systemic source of bias in annotations 
produced by workers. Specifically, many microtask jobs consist of a sequence 
of tasks which share a common format (e.g., label pictures, identify people, 
circle galaxies). Using image-labeling, a canonical microtask job format, we 
discover that earlier tasks shift the distribution of future answers by 
30-50\%, irrespective of the images used. 
Moreover, we show that prior tasks influence 
the content that workers chose to focus on, as well as the richness and 
specialization of their word choice. While these intertask effects can be a 
source of systematic bias, our results suggest that, with appropriate job 
design, it might be possible to leverage these biases to hone worker focus 
and specificity, helping to elicit reproducible, expert-level judgments.

%Faced with the ever-growing scale and complexity of data, academic and 
%industrial researchers are increasingly relying on \textit{human computation} 
%to extract useful knowledge.  Microtask
%platforms, like Amazon Mechanical Turk, act like a human compute cluster,
%making it possible to access human cognition quickly and inexpensively. 
%This has lead to their broad adoption by social and computer scientists.  
%Human computation is being used experimentally in applications
%where expert opinion would normally be sought, such as in reviewing 
%medical images.
%Here we show that the repeated performance of similar tasks, an inherent
%feature of microtask and crowdsourcing platforms, acts as a strong
%and largely unexplored source of bias.  The effects that earlier tasks exert
%on later ones are more severe than effects caused by framing, 
%such as by disclosing a task's purpose.  
%Using image-labeling, the canonical microtask, we show that prior 
%tasks influence the content that workers chose to focus on, as well as the 
%richness and specialization of their vocabulary.    
%While intertask effects can be a source of systematic bias, our
%results suggest that with appropriate task-engineering, it might be possible
%to leverage intertask effects to hone worker focus and specificity, helping
%to elicit reproducible, expert-level judgments.
\end{sciabstract}

@@

Microtask platforms enable a workflow in which workers are like 
input-output devices.  This provides much of the flexibility and 
cost-savings of fully automating the work in a computer program: 
the workforce is available on demand over the Internet using automated 
scripts, and  no contracts or interviews are required.  Many tasks can be
can be performed at a fraction of the cost of recruiting temporary workers.

@@

Given their increasing prevalence, there
is a desire to understand the factors affecting the reliability of 
microtask-based methods.  Researchers have investigated task-design 
parameters such as the 
level of pay \cite{kazai2013analysis}, the training and pre-screening of 
workers \cite{le2010ensuring,paolacci2010running}, and user-interface 
design \cite{Finnerty2013}.  
Researchers have also investigated the effects of \textit{framing}, 
by altering the description of the workflow context 
\cite{Kinnaird2012281}, the purpose of tasks 
\cite{chandler2013breaking}, and the problem description
\cite{thibodeau2013natural}.

Here we focus on a ubiquitous feature of the microtask setup: the 
tendency for microtask workers to perform many tasks in quick succession. 
Even if a worker performs the minimum possible amount of work, 
she will likely have performed several microtasks, since they are typically 
distributed to workers in bundles.  For example, in one bundle, a worker 
might be assigned a set of ten images, and asked to provide descriptive words 
(or \textit{labels}) for each.  On one popular microtask platforms,
Amazon's Mechanical Turk (MTurk), these bundles are called Human 
Intelligence Tasks (HITs), but for the sake of clarity, we will use 
``task'' to mean the smallest repeatable unit of work, which, in our study, 
consists of labeling a single image.

@@

Researchers investigating a somewhat similar class of psychological phenomena,
found that ratings that microtask workers give to items in a list
depend on the order in which the list is presented, an effect known as
\textit{positional bias} \cite{lerman2014leveraging}.  This serves as an 
example of a basic psychological phenomenon impacting crowdsourced data,
and raises questions about what other cognitive biases might
affect microtask work, and how prevelant induced bias might be.

Here we investigate the potential for task-repetition to induce bias in 
microtask-based work.  If such bias is present, it would impact virtually 
every microtask and crowdsourced workflow.
The results we present here show that a worker's response during a task
does indeed depend strongly on the tasks performed beforehand.   
We call these effects that earlier tasks exert on later ones 
\textit{intertask effects}.

As we have described, microtask platforms have been widely adopted as a 
tool, but computer scientists also hold such platforms as the object of 
research, viewing  microtask platforms as a new kind of computing 
architecture.  In analogy to CPUs (central 
processing units) and GPUs (graphics processing units), 
Davis \textit{et al} invoked the \textit{human processing unit}, or HPU \cite{5543192}.
Researchers sought to define a basic HPU instruction set, and
libraries for HPU programming are under active development
 \cite{little2010turkit,minder2011crowdlang,minder2012crowdlang,kittur2011crowdforge}.  
HPU outputs are inherently stochastic, but our revelation of strong intertask 
effects shows that HPUs are subject to \textit{hysteresis}, 
meaning that their 
outputs depend on the history of their inputs.  Reproducibility and 
path-independance are crucial properties of the components in a 
computing architecture, so it to maintaning the view of the HPU computing 
architecture, necessitates a better understanding of the computational 
ramifications of many psychological phenomena.

We investigated intertask effects using image-labeling tasks on the Amazon 
Mechanical Turk (MTurk) microtask platform.  
In addition to being one of the most common kinds of tasks
\cite{chandler2013breaking,Berinsky2012351,Finnerty2013,paolacci2010running},
image-labeling is prototypical for any kind of task in which the worker is 
asked to provide a qualitative free-text response to a prompt.  
This includes such examples as describing
video and audio recordings, summarizing texts, or providing personal 
experiential feedback to a product review.
%in the long run, because image labeling is still very difficult to automate
%\cite{5543192}, and is used in many applications, such as website moderation, 
%computer vision research, and social media analytics (\# insert other refs) 
%\cite{5543192}.

In our first experiment, which we will refer to as 
\textit{intertask-food-objects}, workers were assigned to either a
\textit{food} or \textit{objects} treatment.  Workers in these treatments
performed a set of five \textit{initial tasks}, which each involved 
providing five descriptive labels for an image depicting 
either food or (non-food) objects, depending on the worker's 
treatment (see Fig.~\ref{fig:task}A and B).  Following the initial 
tasks, workers performed five \textit{test tasks}, which were identical for 
both treatments.  The test tasks contained images depicting a 
mixture of food and objects (see Fig.~\ref{fig:task}C), and workers
were again asked to provide five labels.  Five initial 
tasks, together with the test tasks, comprised one HIT\footnote{
	Workers were only allowed to participate in one HIT of a single treatment 
	of a single experiment.
}. During the HIT, workers 
were presented images one at a time, and, crucially, we made no 
distinction between the initial and test tasks.

@@

While this affirms the \textit{existence} of intertask effects, we wish to
understand the nature and severity of the bias induced by prior task 
exposures.  To this end, we define the extent of bias, $\theta$, in terms of
the strength of influence that a perturbation has on worker's responses.
Since different workers generally provide different responses for the same 
task, we characterize worker's responses by a probability distribution.  
The extent of bias is then defined by the difference in the response 
distribution for different treatments of workers\footnote{
	This is a general divergence metric between probability 
	distributions known as the \textit{total variation distance.}
}:
\begin{equation}
	\theta = \frac{1}{2}\sum_{x \in X} \left| p_0(x) - p_1(x) \right|,
	\label{eq:theta}
\end{equation}
where $p_i(x)$ is the probability that a worker subjected to treatment $i$
provides response $x$, from the set of possible responses $X$.
It is important to emphasize that a significant result from a $\chi^2$ test 
only allows us to conclude that $\theta$ is non-zero 
(see the supplementary text).  
Determining $\theta$ from finite samples of 
responses, using purely statistical approaches, is difficult (\#), which 
leads us to reformulate the
problem in terms of classification: given a worker's 
response, can a machine reliably infer what her prior task-exposure had 
been?  Intuitively, the greater the effect that prior tasks have on subsequent
responses, the more accurate the classifier can be.  Formally (as shown
in the supplementary text), the accuracy of such a classifier, $\eta$,
yields a lower bound on the bias:
\begin{equation}
	\theta \geq 2\eta - 1.
	\label{l1}
\end{equation}
By measuring the accuracy of a classifier, we can bound the 
bias induced by any perturbation, including both intertask effects and 
framing.  As defined, bias ranges from 0 (meaning the exposures 
do not affect the probabilities of responses), to 1 
(or 100\%) (meaning that under different exposures, workers provide completely
different sets of responses).

@@

Using a na\"ive Bayes classifier\footnote{
	We discuss our choice of classifier and preprocessing methods, 
	and show similar results obtained by using a support vector 
	machine, in the supplementary material.
}, we found that intertask effects 
lead to a 30\% bias between workers from the \textit{food} and 
\textit{objects} treatments.  This represents a substantial potential to 
distort microtask data.

There has been considerable interest in the effects that framing can have
on microtask responses
\cite{Kinnaird2012281,chandler2013breaking,thibodeau2013natural}, so,
as a point of comparison, we performed a similar experiment 
in which we framed the purpose of the tasks.  In the experiment 
\textit{frame-food-objects}, workers were either told that 
the tasks were ``Funded by the laboratory for the visual perception of Food 
and Ingredients'', or ``\ldots of Objects and Tools''.  
Framing did induce changes in the frequencies of 
word usage at significance (as determined by $\chi^2$ test; $p=0.0012$), but 
the \textit{extent} of framing-induced bias we found was statistically 
insignificant ($p =0.37$) (see Fig.~\ref{fig:theta}).  This suggests 
that prior task exposure is a stronger priming modality than framing.

We conducted variants of these experiments using different images, to see 
whether this trend was robust.  In the experiment 
\textit{intertask-food-culture},
workers were either assigned to a \textit{food} or \textit{culture} treatment.
As before, the initial tasks for the \textit{food} treatment contained images 
depicting food (see Fig.~\ref{fig:task}D). 
In the \textit{culture} treatment, 
the initial tasks contained images depicting cultural scenes 
(of dance, sport, or music) (see Fig.~\ref{fig:task}E).  The test tasks, which were identical for both
treatments, included images depicting meals of identifiable cultural 
origin\footnote{
	Though we attempted to choose more globalized foods, the initial tasks 
	in \textit{intertask-food-culture} are arguably still of
	identifiable cultural origin. This would tend to reduce, rather than 
	inflate, the bias observed in our results 
	(we discuss this further in the supplementary text).
}
(see Fig.~\ref{fig:task}F).

Results for this experiment again showed a 
strong bias as a result of intertask effects ($\theta$~about 50\%) 
(see Fig.~\ref{fig:theta}A).  Again, we compared intertask effects to 
framing, by performing an experiment (\textit{frame-food-culture}) using 
the same test tasks as in \textit{intertask-food-culture}, in which we exposed
workers to a statement framing the purpose of the tasks as the recognition 
of either food or culture.  
For this framing experiment, we did not observe a statistically significant 
difference between the labels attributed by the \textit{food} and 
\textit{culture} treatments during test tasks ($p=0.29$) 
(see Fig.~\ref{fig:theta}).

@@

To better understand the nature of 
intertask effects, we investigated the vocabulary
that workers used to label test tasks. A natural expectation is that, during
the test tasks within
a given experiment, the food-primed workers would use food-related words 
more often than their non-food-primed counterparts.  However, this was not 
generally 
the case. In the \textit{intertask-food-culture} experiment, food-primed
workers actually used significantly \textit{fewer} food-related words 
during test tasks\footnote{
	food-related words were identified using the wordnet corpus, 
	augmented with additional words obtained by crawling a recipe website 
	(see supplementary information for details)
} (see Fig.~\ref{fig:specificity}).  This finding
rules out a seemingly-simple idea that workers emphasize
content that has been present in earlier tasks: seeing content in prior tasks
affects the probability that a worker refers to it in subsequent tasks, but 
does not necessarily \textit{increase} the probability.

To deepen our understanding, we investigated workers' lexical richness in 
reference to food, that is, the number of \textit{unique} food-related words
used.  Even if workers provide an abundance of food-related words, there
can be less diversity if the same words are often repeated.  
This could happen, for example, if workers use generic references 
to food.
Both \textit{intertask} experiments showed that food-exposed workers had 
greater lexical richness, in reference to food, than their non-food-primed 
counterparts (as much as 20\% more) (see Fig.~\ref{fig:specificity}).  
The increase in lexical richness in the case of 
\textit{intertask-food-culture} is particularly noteworthy, because in that 
experiment, food-primed workers made fewer total references to food.  
We also observed enrichment of the food-related lexicon in the 
\textit{priming-food-objects}, and \textit{echo-food-objects} experiments, 
although to a lesser extent.

The observations regarding lexical richness lead us to suspect that 
initial tasks might influence workers to use more refined or specialized 
words, when referring to aspects of content that had been present in the 
initial tasks (i.e. food).  This would help explain why, in the case of 
\textit{intertask-food-culture}, we observed a significantly greater 
\textit{diversity} of food-related words despite their significantly lesser 
\textit{total amount}.
To test whether exposure to food in initial tasks increased the specialization
of food-references, we used the wordnet corpus to operationalize the 
notion of word specialization.  Wordnet provides a set of hypernym-hyponym 
relations between 117,798 English nouns (\#).  Hypernyms are generalizations 
(for example, ``bread'' is a hypernym for 
``pumpernickel''), while hyponyms are specializations.
For each experiment, we compared food-related words provided by workers from 
one treatment to those from the other, and tallied the excess number of cases
where one treatment's words were more specialized than the other, as a 
percentage of the total 
(see Fig.~\ref{fig:specificity}B) 
(details for this calculation are in the supplementary text).

In all experiments except \textit{frame-food-culture}, food-primed workers 
used significantly more specialized words than their non-food-primed 
counterparts (about 15\% more) (see Fig.~\ref{fig:specificity}C).
It is interesting that such substantial increases in both the lexical 
richness and specialization of food-related words 
held for \textit{intertask-food-culture}, where, as mentioned, we observed 
that food-exposed workers made \textit{fewer} references to food overall. 
These observations point 
to countervailing factors: one factor tending to activate the more 
specialized and less common food-related words 
(yielding greater lexical richness and specialization), and the other tending 
to suppress certain, presumably more common and generic words 
(yielding fewer food-related words in total).

@@

\section*{Supplementary Material}
\renewcommand{\figurename}{Figure S\!\!}
\renewcommand{\tablename}{Table S\!\!}
\setcounter{figure}{0}
\setcounter{table}{0}

\subsection*{Materials and Methods}

\paragraph{Microtask setup.}
Altogether the experiments presented here consisted of data from 3200 HITs 
(bundles of 10-image labeling tasks).  In all experiments, each treatment
consisted of results from at least 119 workers.  
For the purpose of statistical calculations,
treatments having more than 119 workers were subsampled, producing a uniform
sample size accross all treatments.  The \textit{food} and \textit{objects}
treatments in the \textit{intertask-food-objects} experiment were further
broken down into sub-treatments (each containing at least 119 workers),
corresponding to different orderings of the test tasks.  These alternative
orderings were obtained by rotation of the test tasks 
(i.e. moving all tasks forward by one position, and making the last task 
become the first test task).  This ensured that each test task appeared in
each of the five test-task positions.

The HITs (as defined in the main text) for all treatments were presented as 
a series of slides.  The first 
slide consisted of a brief set of instructions.  In framing-treatments, a 
framing slide was shown immediately after the instructions, see 
Fig. S\ref{fig:hit_preamble}.  In intertask-treatments, a set of initial
tasks were presented (and no framing slide).  Following the
framing slide or initial tasks, the test-tasks were presented.

All initial and test tasks had the format shown in 
Fig.~S\ref{fig:hit_preamble}: an image taking up most of
the task pane, with five text inputs and a ``next'' button.  Images were
presented in this way, one at a time. 
For treatments having initial tasks, the test
tasks followed seemlessly, and no distinction was made between the test 
tasks and the initial tasks.

In an attempt to increase the strength of framing, we included the
\textit{echo-food-objects} treatment, in which, during the framing slide,
the worker was required to echo the frame using a combo-box input.  The
instructions, both kinds of framing slide (simple and echoed), 
and an example task slide, are shown 
in Fig.~S\ref{fig:hit_preamble}.

Given the setup described above, the only tasks that workers in framing 
treatments perform are the test tasks, whereas in the 
intertask-treatments, the test tasks were performed as tasks
5 through 10.  
It could be argued that a better comparison would be achieved
by having all workers perform some initial tasks, and simply ensuring that
the initial tasks are \textit{not} varied between framing treatments. 
We took this approach in the experiment 
\textit{frame-food-culture}, in which workers performed the same 
initial tasks containing images that depicted food in both the \textit{food}
and \textit{culture} treatments.  If anything, this appears
to have reduced the strength of framing for that experiment.

\begin{figure}
	\includegraphics[scale=0.8]{figs/tasks.pdf}
	\caption{Examples of A) instructions; B) a frame, as shown in 
		\textit{frame1:obj}, and similar to those shown in 
		\textit{frame1:food} and \textit{frame2:food}, and 
		\textit{frame2:cult};  
		C) an echoed frame, as used in \textit{echo1:obj}, and similar to that
		used in \textit{echo1:food}; and D) an 
		example of an image-labeling task.
	}
	\label{fig:hit_preamble}
\end{figure}

\paragraph{Selection of images.}

The images used in the initial tasks for \textit{intertask-food-objects}  
are shown in Figs.~S\ref{fig:task1:food} (\textit{food} treatment) and 
S\ref{fig:task1:obj} (\textit{objects} treatment).
We selected images for the initial tasks that show either food or objects, 
according to the treatment in which they were used.  
We took care to exclude objects from
the food-images (except in some cases a dish supporting the food), and 
to exclude food from the object-images.  The images used for the test tasks
of \textit{intertask-food-objects} and \textit{frame-food-objects} are shown 
in Fig.~S\ref{fig:task1:test}.

The images used in the initial tasks
for \textit{intertask-food-culture} are shown in Figs.~S\ref{fig:task2:food} 
(\textit{food} treatment) and S\ref{fig:task2:cult} 
(\textit{culture} treatment).
We selected the initial images for the treatments of this experiment to 
respectively reflect depict food and culture.

Naturally, since food is a very
important aspect of culture, pictorial depictions of the two concepts 
cannot be cleanly separated,
and images depicting food inevitably also depict culture.
The fact that the inital-task images for the food treatment of 
\textit{intertask-food-culture} do also depict culture
will only tend to make them more similar to those from the \textit{culture}
treatment, which would tend to \textit{reduce} the severity of intertask 
effects.  In any case, since we still found strong intertask effects, 
the presence of culture in the food-related initial images of 
\textit{intertask-food-culture} does not seem to have been problematic.
The images used in the test tasks of \textit{intertask-food-culture} and
\textit{frame-food-culture} are shown in Fig.~S\ref{fig:task2:test}.

\begin{figure}
	\includegraphics{figs/task1-food.pdf}
	\caption{
		Images used in the initial tasks for the
		\textit{food} treatment of \textit{intertask-food-objects}.  
		The numbers show the order in which the 
		images were presented to workers.
	}
	\label{fig:task1:food}
\end{figure}

\begin{figure}
	\includegraphics{figs/task1-obj.pdf}
	\caption{
		Images used in the initial tasks for the
		\textit{objects} treatment of \textit{intertask-food-objects}.  
		The numbers show the order in which the images were presented to 
		workers.
	}
	\label{fig:task1:obj}
\end{figure}
